Implementation of a stable, high-power optical lattice for quantum gas microscopy

### 2018-10-14 LatticeOptics/latticeOpticsV4.tex

\author{A.Mazurenko, S. Blatt, F. Huber, M. F. Parsons, C. S. Chiu, G. Ji, D. Greif}
\author{M. Greiner}
\affiliation{
  Department of Physics, Harvard University,
  Cambridge, Massachusetts, 02138, USA}

\date{\today}
\begin{abstract}
 We describe the design of a stable high power $1064\,$nm laser system for practical use in optical lattice experiments. The system is based on a low noise laser amplified by an array of four heavily modified, high-power fiber amplifiers. The beam intensity is stabilized with a nonlinear feedback loop. The position of the produced optical lattice is monitored in real time and its stability is characterized.
\end{abstract}
\maketitle

Optical lattices holding an array of ultracold atoms have become a powerful experimental platform for a wide variety of investigations into various forms of quantum simulation \cite{Friedenauer2008, Kim2010, Struck2011, Simon2011, Yan2013, Drewes2016, Murmann2015}. 
The recent advent of quantum gas microscopy~\cite{Bakr2009, Sherson2010, Haller2015, Cheuk2015, Parsons2015, Edge2015, Omran2015, Greif2016, Cheuk2016, Parsons2016, Boll2016, Cheuk2016a, Brown2016} has enabled studies that exhibit unprecedented control over individual atoms at extremely low energy scales, particularly those governed by the Hubbard Hamiltonian ~\cite{Hubbard1963}.
However, this progress has placed ever more stringent technical constraints on the lasers used to trap and manipulate the atomic systems.
For example, heating due to the intensity noise cannot be allowed to approach the longest thermalization timescales, lest the system fail to thermalize~\cite{Savard1997}.
Even if the system thermalizes, heating rates can limit the interrogation time available to probe a low-temperature state.
Similarly, experiments that rely on a small number of atoms held by multiple traps place stringent requirements on the relative positional stability of the traps, to ensure that the operations performed by the traps have adequate fidelity.

In this work we discuss a high power optical lattice system that exhibits levels of intensity noise below $-120\,$dBc between $1\,$kHz and $3\,$MHz.
The stability of the lattice position is below the single site level and the stability of the harmonic confinement is at the few site level.

@@

The coarse architecture of the system is shown in figure \ref{fig:high_level} and is based on a single low-power, low-noise seed laser that supplies light to an array of four high-power fiber amplifiers (one for a transport beam~\cite{Huber2014}, and one each for the $X,Y,Z$ lattices~\cite{Parsons2016}).
These amplifiers generate enough light to perform site resolved imaging~\cite{Parsons2015}, but also substantially increase the level of intensity and pointing noise. 
The fiber amplifier output is controlled by two feedback loops, which stabilize the light in low power and high power regimes, discussed later.
After intensity stabilization, the laser is mode-shaped to a desired waist.
The optical system is engineered to be passively stable in terms of pointing, but the position and phase of the optical lattice are monitored in real time, and beam pointing can be controlled with remotely actuated mirrors.

@@

The NuAmp suffers from several technical problems. 
It is controlled via a USB interface, which complicates communications with an array of these devices.  
Worse yet, it is suspected of causing ground loops which add noise and interference to the analog systems. 
This added noise can be seen in figure \ref{fig:noises}, where the fiber-amplified light exhibits many more spurs in the intensity noise spectrum than the seed laser.
We solve this problem by replacing the USB control board with a custom-designed solution  that is compatible with our custom control system and engineered to avoid ground loops.

@@

These upgrades serve to suppress many of the noise spurs measured in the relative intensity noise (RIN) of the fiber amplifier. 
This can be as seen by comparing figure \ref{fig:noises} (a) and (b), particularly in the region of approximately $10$ kHz. 
The measurements in parts (a)-(d) were carried out on a single fiber amplifier, prior to and after the modifications. The initial and final noise spectra were verified to be consistent between amplifiers coming from the same manufacturing batch.
The figure is based on fiber amplifiers from a relatively new batch, which, even in its unmodified configuration does not exhibit a large switching supply spur. 
This is not the case for certain older batches, where the spur is approximately 30 dB higher than the background.

The SBS noise is fundamental to the fiber amplifier architecture and cannot be entirely avoided. However, since it is a function of fiber length and optical power, it can be minimized. To this end, we use custom-ordered amplifiers with a $0.5\,$m length output fiber, compared to the few meter standard lengths.
In addition, we set the lowest pump diode current that yields enough power to perform imaging. 
The rise of the noise floor as a function of the pump diode current can be seen in figure \ref{fig:noises}(a).

@@

In order to address these challenges, we use a carefully engineered the optical system shown in figure \ref{fig:optical_layout}.
This system collimates and polarizes the beam, controls its intensity and ensures good pointing stability.
Due to the high laser power, fused silica glass and IBS coatings are used wherever possible.
Since reflections pose a danger to the users and equipment, all undesired beams above $100\,$ mW are directed to water cooled beam-dumps (to prevent temperature gradients on the table) placed as far as practical from critical beam paths.
Weaker beams are dumped on uncooled diffusive beam catchers.

@@

After the isolation stage, we use the sequence of optical elements shown in figure \ref{fig:optical_layout}, consisting of a polarizer, half-wave plate (HWP), Berek compensator mounted on a precision galvo (Thorlabs GVS002, or Camtech 8320K, depending on beam path) half-wave plate, and polarizer. 
The Berek compensator is a $z$-cut quartz plate coated to be anti-reflecting at $1064\,$nm. 
By tilting the plate about its vertical axis by a few degrees, the extraordinary axis is mixed into the propagation of the beam, leading to  tilt-dependent birefringence that can be tuned from a zero-wave plate past a half-wave plate, represented by a retardation angle $\phi$.
Combined with the subsequent polarization optics, rotation of the Berek compensator can vary the transmitted power with perfect contrast. 
Figure \ref{fig:berek} shows the transmitted power as a function of retrardation angle $\phi$ and fast-axis angle $\theta$ of the waveplates surrounding the compensator.
\begin{figure}
  \begin{center}
    \includegraphics{fig/berek_sim.pdf}
    \caption{\textbf{Berek compensator} transmission as a function of the Berek tilt angle $\phi$ and waveplate tilt angles.}\label{fig:berek}
  \end{center}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{fig/step_response.pdf}
    \caption{\textbf{Berek compensator based feedback loop}. This Berek compensator based feedback loop is used to vary the beam intensity at relatively high power. (a) Shows the small signal impulse response and (b-c) show the corresponding Bode plots.}\label{fig:berek_step_response}
  \end{center}
\end{figure}
A key point, as seen in figure \ref{fig:berek}, is that if the waveplates are rotated together, the maximum transmitted power is fixed, while the minimum can be raised.
The waveplates are detuned such that the maximum power needed in \textit{interaction} mode is the minimum power transmitted by the Berek compensator system, strongly suppressing fluctuations due to angle positioning noise of the Berek compensator. 
Naturally, rotation of a parallel plate in the beam path induces a shift of the beam, but this is acceptable because the Berek compensator is only active during the imaging phase of the experiment, when the position of the underlying harmonic trap is largely irrelevant.
Fortunately, the spatial phase of the optical lattice is set by the retro-reflector mirror, which does not move, and is not affected by the rotation of the Berek compensator.
This system functions as the actuator arm of a digital feedback loop, where the power of the beam is measured using a low noise photodiode. This feedback loop regulates the lattice between $1$ W of optical power and its maximum value (which depends on the amplifier, but is $\approx20$ W).
The step response and corresponding Bode plots~\cite{Bechhoefer2005} are shown in figure \ref{fig:berek_step_response}.
A polarization rotation by $2 \pi$ is produced by approximately $3^\circ$ of axial rotation of the quartz plate. 
Since we never need to rotate the plate outside this range, the position is constrained to this window in software.
Since two feedback loops are present in the system (the second one detailed in the next section), care must be taken so that they do not function simultaneously. 
To accomplish this, the low power loop is always manually railed to its high extreme when the high power loop is active, and the digital high power loop is disabled when the low power loop is active.

@@

\begin{figure}
  \begin{center}
    \includegraphics{fig/circuits.pdf}
    \caption{\textbf{Feedback circuits.} (a) The reverse-biased photodiode. (b) Externally referenced phase-locked loop. (c) Nonlinear loop filter. (d) RF power modulation using a mixer.}\label{fig:circuits}
  \end{center}
\end{figure}
The feedback circuit is shown in figure \ref{fig:circuits}a. As seen in figure \ref{fig:optical_layout}, a pair of low noise photodiodes measures the optical power of the beam sampled with a weakly reflective optic (Thorlabs BSF10-C). 
The beam sampler is wedged in such a way that the two reflections diverge from one another, allowing us to illuminate both photodiodes. 
One of the sampled beams is attenuated by an additional factor of $20\times$ by reflection off of an uncoated fused silica glass plate, such that the dynamic range of its photodiode covers $0-20$ W of the beam directed at the atoms, and controls the Berek compensator based, high-power feedback loop. 
The other beam is not attenuated, save by the beam sampler, and its dynamic range covers $0-1$ W.
To prevent damage to the low power photodiode when the system is at its highest powers, a shutter (Stanford Research Systems SR475) blocks the light when the high power loop is in use.
The low power photodiode passes its signal $V_{\mathrm{pd}}$ to a loop filter circuit, which compares it to the control system's set-point, generated by digital to analog converter.
The loop filter's output controls the IF port of a microwave mixer (Minicircuits ZFM-2-S+), effectively modulating the RF power supplied by a low noise phase-locked loop (PLL).
The resulting RF signal is then amplified and sent to the AOM.
Let us now consider every part of this signal chain individually.

@@

The chosen photodiode (Hamamatsu G8370-01) is an InGaS based, small-area, low-noise design with a high efficiency at 1064 nm.
Prior to use, the protective window is removed from the photodiode to prevent interference effects, as seen in figure \ref{fig:cut_pd}.
Further, to reduce the capacitance of the system, a reverse bias of $6.9\,$V is applied to the photodiode by a temperature-stabilized voltage reference (Linear Technology LM399).
The trans-impedance circuit is shown in figure \ref{fig:circuits}. 
It converts the photo-current supplied by the photodiode to a voltage signal using the low noise Texas Instruments OPA843 op-amp. 
To reduce electrical pickup and prevent ground loops, the electronic assembly is housed in a metal enclosure and mounted to the optical table using electrically insulating mica posts.

@@

\subsubsection{Digital to analog converter}
The digital to analog converter (DAC, Texas Instruments DAC9881) connects the experimental digital control system to the analog control loop for the lattice. 
By definition, an ideal control loop would translate any noise on the analog signal from the DAC to the laser light, meaning that a poorly designed DAC would inherently limit the lattice laser performance.
The noise on the DAC, when it is programmed to output a constant voltage is shown in figure \ref{fig:da_rin}, and is found to be roughly constant at a level of $-150$ dBc - more than acceptable.
When the state of the DAC is changed, such as during a ramp, this noise level may be higher, due to the so-called ``glitch'' energy, as seen in the figure.
The transient noise, while unfortunate, is mitigated by two factors: (1) it is only present when the lattice power is changing, which does not happen for most of the sequence, and a lattice at fixed power will have negligible RIN contributions. Further, (2) the loop filter eliminates the noise at higher frequencies.

@@

The connection between the actuation and monitoring mechanisms of the feedback system is the loop filter, which compares the \textit{set-point} signal, $v_s(t)$, supplied by the experimental control system and the monitoring signals $v_m(t)$, to adjust the actuation accordingly.
We use a simple, purely integrating design based in three stages, as seen in figure \ref{fig:circuits}, based on TI's OPA211 op-amp.
The first stage computes and amplifies an error signal, $e(t)=v_m(t)-v_s(t)$, which is integrated by the second stage. 
The integral gain is set using a low-drift trim-pot. 
This is the case because within relevant frequency ranges, the AOM can be coarsely approximated by a pure time delay process (with delay given by the RF travel time across the crystal $\approx 1 \mu\mathrm{s}$).
In this scenario, the $P$ term must be small to avoid instability at the $\pi$ phase shift point, and a pure $I$ controller is optimal~\cite{Skogestad2001}.
For debugging purposes, the loop filter features buffered photodiode monitor and error signal monitor outputs.

The third stage of the loop filter is used to accommodate the fact that the actuation arm is nonlinear - the optical power transmitted by the AOM is proportional to the RF power incident on it $P_{\mathrm{rf}}\propto v_{if}^2$, where $v_{if}$ is the actuation signal supplied to the IF port of the mixer.
In effect, if the rest of the feedback system was linear, this would correspond to a quadratically increasing gain until the AOM is saturated, which could lead to slow behavior at low signal levels and ringing at higher signal levels.
Even worse, since the mixer is insensitive to the sign of the voltage applied, the gain is actually negative for negative voltages, resulting in an unstable system, which thus requires a current clamp.
This stage linearizes the response by increasing the gain at low signal levels and suppressing it at high signal levels by using a diode (Digi-Key BAT54S) in the feedback arm.
This can be seen by considering two extreme cases - far below the diode drop, the diode is an open circuit and the gain is set by $R_l$, where above the diode drop it is a closed circuit where the gain is set by the resistor $R_h$. 
Naturally, the voltage drop over the diode is chosen such that it is approximately half the maximum desired signal level of the photodiode ($3.3\,$V).
Although a large improvement, the bandwidth still varies over the dynamic range of the loop, but only by approximately a factor of two, as seen in figure \ref{fig:bandwidth}.
Were it not for the linearization, the gain of the mixer would change by a factor of $10$ over a dynamic range spanning two orders of magnitude.
Another function of the last stage is to clamp the voltage to be strictly $>0$, in order to prevent an unstable positive feedback behavior, which is done using diode clamp on the output arm~\cite{Horowitz2015}.
Last of all, since the mixer is a current driven device, the last stage of the feedback loop uses TI's OPA627 op-amp which is able to supply tens of mA of current at an output impedance of $50\,\Omega$.

@@

Although it is an essential part of the system, the loop filter is also one of its major limitations. Figure \ref{fig:noises} shows the noise performance of the output of the fiber amplifiers in open loop as compared to the noise in closed loop. Although noise near DC is lowered (as expected from a closed loop system), noise outside the loop bandwidth is increased in the range of $10$ to $450$ kHz as seen in subfigures (c) and (d).
This can be explained by the Johnson noise~\cite{Horowitz2015} on the $1$ k$\Omega$ input resistors, amplified by the amplification stages of the loop filter.
The sharp drop off at the $450$ kHz point is due to a low-pass filter placed on the output of the loop filter, far above the maximum bandwidth set by the AOM.
When this was determined to be a problem, we decided that a bandwidth of higher than $10\,$ kHz was not required in the foreseeable future. 
As a result, we added a first order low pass with a $3$ dB frequency of $10$ kHz leading to the improvement shown in figure \ref{fig:low_pass_noises} (the fiber amplifier used to take this data was different from the one used for \ref{fig:noises}, although their performance is comparable.

@@

The absolute positions of lattice sites cannot be allowed to change by more than a small fraction ($\approx 10\%$) from experimental shot to shot.
This is ensured via careful design of the optical assembly, since the phase of a retro-reflected lattice is set by the retro-reflecting mirror, which is placed approximately $3\,$cm from the atoms~\cite{Huber2014}. 
We have characterized the phase stability of the optical lattice by looking at a long sequence of densely populated images, and found that the phase varies by much less than one lattice site over the course of hours, as seen in figure \ref{fig:averaged_mott}.

@@

The position of the atomic cloud, however, is not determined by the phase of the lattice, but by the underlying harmonic confinement. 
In experiments where the entire cloud is addressed by beams auxiliary to the lattice (such as in \cite{Mazurenko2016} or \cite{Choi2016}), drift or fluctuation of the lattice position is important.

@@

Over the course of 250 sequential experimental realizations, we monitored the position of the beam on the camera and the position of the atoms. 
First, we characterized the passive stability of the optical axis along the transverse $x_t$ and longitudinal $x_l$ directions. 
Figure \ref{fig:stability} a shows a histogram of displacements of atoms between subsequent experimental realizations, showing that the RMS displacement is $0.54(3)$($0.98(4)$) sites along the $x_t$($x_l$) directions.
The difference between displacements along is a geometric effect consistent with the mechanical assembly of the experiment~\cite{Huber2014}.
The distributions are centered within $1\sigma$ of zero along both directions ($-.03(4)$ and $0.01(5)$ sites for transverse and longitudinal, respectively).
If the trap followed a random walk with these step sizes, the experiment would be in trouble, since after only $100$ experimental shots, the trap would move by more than $10$ lattice sites.
The situation is more fortunate, however, since subsequent steps are correlated, as shown in figure \ref{fig:stability} (b).
Once this correlation is taken into account, we estimate a drift of $8(1)\times 10^{-3}\mathrm{sites}/\mathrm{realization}$ in the transverse direction and $12(2)\times 10^{-3}\mathrm{sites}/\mathrm{realization}$ in the axial direction, compared to a beam waist of approximately $80$ $\mu$m ($\approx 140$ sites).

@@

In a real experiment, we cannot directly measure the position of the atoms and must rely on the measurement performed by the camera, related by an affine transformation.
To determine the parameters of the affine transform, we use the remote actuators to purposefully misalign the lattice beams for a small number of shots, and simultaneously measure the atomic positions and the camera reading, then perform a least-squares fit, shown in figure \ref{fig:stability} (c).

>>>>

\author{A. Mazurenko, S. Blatt, F. Huber, M. F. Parsons, C. S. Chiu, G. Ji, D. Greif}
\author{M. Greiner}
\affiliation{
  Department of Physics, Harvard University,
  Cambridge, Massachusetts, 02138, USA}

\date{\today}
\begin{abstract}
 We describe the design of a stable high power 1064~nm laser system for practical use in optical lattice experiments. The system is based on a low noise laser amplified by an array of four heavily modified, high-power fiber amplifiers. The beam intensity is stabilized with a nonlinear feedback loop. The position of the produced optical lattice is monitored in real time and its stability is characterized.
\end{abstract}
\maketitle

Optical lattices holding an array of ultracold atoms have become a powerful experimental platform for a wide variety of investigations into various forms of quantum simulation~\cite{Friedenauer2008, Kim2010, Struck2011, Simon2011, Yan2013, Drewes2016, Murmann2015}.
The recent advent of quantum gas microscopy~\cite{Bakr2009, Sherson2010, Haller2015, Cheuk2015, Parsons2015, Edge2015, Omran2015, Greif2016, Cheuk2016, Parsons2016, Boll2016, Cheuk2016a, Brown2016} has enabled studies that exhibit unprecedented control over individual atoms at extremely low energy scales, particularly those governed by the Hubbard Hamiltonian ~\cite{Hubbard1963}.
However, this progress has placed ever more stringent technical constraints on the lasers used to trap and manipulate the atomic systems.
For example, heating due to the intensity noise cannot be allowed to approach the longest thermalization timescales, lest the system fail to thermalize~\cite{Savard1997}.
Even if the system thermalizes, heating rates can limit the interrogation time available to probe a low-temperature state.
Similarly, experiments that rely on a small number of atoms held by multiple traps place stringent requirements on the relative positional stability of the traps, to ensure that the operations performed by the traps have adequate fidelity.

In this work we discuss a high power optical lattice system that exhibits levels of intensity noise below $-120\,$dBc/Hz between $1\,$kHz and $3\,$MHz Fourier frequency.
The stability of the lattice position is below the single site level and the stability of the harmonic confinement is at the few site level.

@@

The coarse architecture of the system is shown in Fig.~\ref{fig:high_level} and is based on a single low-power, low-noise seed laser that supplies light to an array of four high-power fiber amplifiers (one for a transport beam~\cite{Huber2014}, and one each for the $X,Y,Z$ lattices~\cite{Parsons2016}).
These amplifiers generate enough light to perform site resolved imaging~\cite{Parsons2015}, but also substantially increase the level of intensity and pointing noise.
The fiber amplifier output is controlled by two feedback loops, which stabilize the light in low power and high power regimes, discussed later.
After intensity stabilization, the laser is mode-shaped to a desired waist.
The optical system is engineered to be passively stable in terms of pointing, but the position and phase of the optical lattice are monitored in real time, and beam pointing can be controlled with remotely actuated mirrors.

@@

The NuAmp suffers from several technical problems.
It is controlled via a USB interface, which complicates communications with an array of these devices.
Worse yet, it is suspected of causing ground loops which add noise and interference to the analog systems.
This added noise can be seen in Fig.~\ref{fig:noises}, where the fiber-amplified light exhibits many more spurs in the intensity noise spectrum than the seed laser.
We solve this problem by replacing the USB control board with a custom-designed solution  that is compatible with our custom control system and engineered to avoid ground loops.

@@

These upgrades serve to suppress many of the noise spurs measured in the relative intensity noise (RIN) of the fiber amplifier.
This can be as seen by comparing Fig.~\ref{fig:noises} (a) and (b), particularly in the region of approximately $10$ kHz.
The measurements in parts (a)-(d) were carried out on a single fiber amplifier, prior to and after the modifications. The initial and final noise spectra were verified to be consistent between amplifiers coming from the same manufacturing batch.
The figure is based on fiber amplifiers from a relatively new batch, which, even in its unmodified configuration does not exhibit a large switching supply spur.
This is not the case for certain older batches, where the spur is approximately 30 dB higher than the background.

The SBS noise is fundamental to the fiber amplifier architecture and cannot be entirely avoided. However, since it is a function of fiber length and optical power, it can be minimized. To this end, we use custom-ordered amplifiers with a $0.5\,$m length output fiber, compared to the few meter standard lengths.
In addition, we set the lowest pump diode current that yields enough power to perform imaging.
The rise of the noise floor as a function of the pump diode current can be seen in Fig.~\ref{fig:noises}(a).

@@

In order to address these challenges, we use a carefully engineered the optical system shown in Fig.~\ref{fig:optical_layout}.
This system collimates and polarizes the beam, controls its intensity and ensures good pointing stability.
Due to the high laser power, fused silica glass and IBS coatings are used wherever possible.
Since reflections pose a danger to the users and equipment, all undesired beams above $100\,$ mW are directed to water cooled beam-dumps (to prevent temperature gradients on the table) placed as far as practical from critical beam paths.
Weaker beams are dumped on uncooled diffusive beam catchers.

@@

After the isolation stage, we use the sequence of optical elements shown in Fig.~\ref{fig:optical_layout}, consisting of a polarizer, half-wave plate (HWP), Berek compensator mounted on a precision galvo (Thorlabs GVS002, or Camtech 8320K, depending on beam path) half-wave plate, and polarizer.
The Berek compensator is a $z$-cut quartz plate coated to be anti-reflecting at $1064\,$nm.
By tilting the plate about its vertical axis by a few degrees, the extraordinary axis is mixed into the propagation of the beam, leading to  tilt-dependent birefringence that can be tuned from a zero-wave plate past a half-wave plate, represented by a retardation angle $\phi$.
Combined with the subsequent polarization optics, rotation of the Berek compensator can vary the transmitted power with perfect contrast.
Figure~\ref{fig:berek} shows the transmitted power as a function of retrardation angle $\phi$ and fast-axis angle $\theta$ of the waveplates surrounding the compensator.
\begin{figure}
  \begin{center}
    \includegraphics{fig/berek_sim.pdf}
    \caption{\textbf{Berek compensator} transmission as a function of the Berek tilt angle $\phi$ and waveplate tilt angles.}\label{fig:berek}
  \end{center}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{fig/step_response.pdf}
    \caption{\textbf{Berek compensator based feedback loop}. This Berek compensator based feedback loop is used to vary the beam intensity at relatively high power. (a) Shows the small signal impulse response and (b-c) show the corresponding Bode plots.}\label{fig:berek_step_response}
  \end{center}
\end{figure}
A key point, as seen in Fig.~\ref{fig:berek}, is that if the waveplates are rotated together, the maximum transmitted power is fixed, while the minimum can be raised.
The waveplates are detuned such that the maximum power needed in \textit{interaction} mode is the minimum power transmitted by the Berek compensator system, strongly suppressing fluctuations due to angle positioning noise of the Berek compensator.
Naturally, rotation of a parallel plate in the beam path induces a shift of the beam, but this is acceptable because the Berek compensator is only active during the imaging phase of the experiment, when the position of the underlying harmonic trap is largely irrelevant.
Fortunately, the spatial phase of the optical lattice is set by the retro-reflector mirror, which does not move, and is not affected by the rotation of the Berek compensator.
This system functions as the actuator arm of a digital feedback loop, where the power of the beam is measured using a low noise photodiode. This feedback loop regulates the lattice between $1$ W of optical power and its maximum value (which depends on the amplifier, but is $\approx20$ W).
The step response and corresponding Bode plots~\cite{Bechhoefer2005} are shown in Fig.~\ref{fig:berek_step_response}.
A polarization rotation by $2 \pi$ is produced by approximately $3^\circ$ of axial rotation of the quartz plate.
Since we never need to rotate the plate outside this range, the position is constrained to this window in software.
Since two feedback loops are present in the system (the second one detailed in the next section), care must be taken so that they do not function simultaneously.
To accomplish this, the low power loop is always manually railed to its high extreme when the high power loop is active, and the digital high power loop is disabled when the low power loop is active.

@@

\begin{figure}
  \begin{center}
    \includegraphics{fig/circuits.pdf}
    \caption{\textbf{Feedback circuits.} (a) The reverse-biased photodiode. (b) Externally referenced phase-locked loop. (c) Nonlinear loop filter. (d) RF power modulation using a mixer.}\label{fig:circuits}
  \end{center}
\end{figure}
The feedback circuit is shown in Fig.~\ref{fig:circuits}a. As seen in Fig.~\ref{fig:optical_layout}, a pair of low noise photodiodes measures the optical power of the beam sampled with a weakly reflective optic (Thorlabs BSF10-C).
The beam sampler is wedged in such a way that the two reflections diverge from one another, allowing us to illuminate both photodiodes.
One of the sampled beams is attenuated by an additional factor of $20\times$ by reflection off of an uncoated fused silica glass plate, such that the dynamic range of its photodiode covers $0-20$ W of the beam directed at the atoms, and controls the Berek compensator based, high-power feedback loop.
The other beam is not attenuated, save by the beam sampler, and its dynamic range covers $0-1$ W.
To prevent damage to the low power photodiode when the system is at its highest powers, a shutter (Stanford Research Systems SR475) blocks the light when the high power loop is in use.
The low power photodiode passes its signal $V_{\mathrm{pd}}$ to a loop filter circuit, which compares it to the control system's set-point, generated by digital to analog converter.
The loop filter's output controls the IF port of a microwave mixer (Minicircuits ZFM-2-S+), effectively modulating the RF power supplied by a low noise phase-locked loop (PLL).
The resulting RF signal is then amplified and sent to the AOM.
Let us now consider every part of this signal chain individually.

@@

The chosen photodiode (Hamamatsu G8370-01) is an InGaS based, small-area, low-noise design with a high efficiency at 1064 nm.
Prior to use, the protective window is removed from the photodiode to prevent interference effects, as seen in Fig.~\ref{fig:cut_pd}.
Further, to reduce the capacitance of the system, a reverse bias of $6.9\,$V is applied to the photodiode by a temperature-stabilized voltage reference (Linear Technology LM399).
The trans-impedance circuit is shown in Fig.~\ref{fig:circuits}.
It converts the photo-current supplied by the photodiode to a voltage signal using the low noise Texas Instruments OPA843 op-amp.
To reduce electrical pickup and prevent ground loops, the electronic assembly is housed in a metal enclosure and mounted to the optical table using electrically insulating mica posts.

@@

\subsubsection{Digital to analog converter}
The digital to analog converter (DAC, Texas Instruments DAC9881) connects the experimental digital control system to the analog control loop for the lattice.
By definition, an ideal control loop would translate any noise on the analog signal from the DAC to the laser light, meaning that a poorly designed DAC would inherently limit the lattice laser performance.
The noise on the DAC, when it is programmed to output a constant voltage is shown in Fig.~\ref{fig:da_rin}, and is found to be roughly constant at a level of $-150$ dBc - more than acceptable.
When the state of the DAC is changed, such as during a ramp, this noise level may be higher, due to the so-called ``glitch'' energy, as seen in the figure.
The transient noise, while unfortunate, is mitigated by two factors: (1) it is only present when the lattice power is changing, which does not happen for most of the sequence, and a lattice at fixed power will have negligible RIN contributions. Further, (2) the loop filter eliminates the noise at higher frequencies.

@@

The connection between the actuation and monitoring mechanisms of the feedback system is the loop filter, which compares the \textit{set-point} signal, $v_s(t)$, supplied by the experimental control system and the monitoring signals $v_m(t)$, to adjust the actuation accordingly.
We use a simple, purely integrating design based in three stages, as seen in Fig.~\ref{fig:circuits}, based on TI's OPA211 op-amp.
The first stage computes and amplifies an error signal, $e(t)=v_m(t)-v_s(t)$, which is integrated by the second stage.
The integral gain is set using a low-drift trim-pot.
This is the case because within relevant frequency ranges, the AOM can be coarsely approximated by a pure time delay process (with delay given by the RF travel time across the crystal $\approx 1 \mu\mathrm{s}$).
In this scenario, the $P$ term must be small to avoid instability at the $\pi$ phase shift point, and a pure $I$ controller is optimal~\cite{Skogestad2001}.
For debugging purposes, the loop filter features buffered photodiode monitor and error signal monitor outputs.

The third stage of the loop filter is used to accommodate the fact that the actuation arm is nonlinear - the optical power transmitted by the AOM is proportional to the RF power incident on it $P_{\mathrm{rf}}\propto v_{if}^2$, where $v_{if}$ is the actuation signal supplied to the IF port of the mixer.
In effect, if the rest of the feedback system was linear, this would correspond to a quadratically increasing gain until the AOM is saturated, which could lead to slow behavior at low signal levels and ringing at higher signal levels.
Even worse, since the mixer is insensitive to the sign of the voltage applied, the gain is actually negative for negative voltages, resulting in an unstable system, which thus requires a current clamp.
This stage linearizes the response by increasing the gain at low signal levels and suppressing it at high signal levels by using a diode (Digi-Key BAT54S) in the feedback arm.
This can be seen by considering two extreme cases - far below the diode drop, the diode is an open circuit and the gain is set by $R_l$, where above the diode drop it is a closed circuit where the gain is set by the resistor $R_h$.
Naturally, the voltage drop over the diode is chosen such that it is approximately half the maximum desired signal level of the photodiode ($3.3\,$V).
Although a large improvement, the bandwidth still varies over the dynamic range of the loop, but only by approximately a factor of two, as seen in Fig.~\ref{fig:bandwidth}.
Were it not for the linearization, the gain of the mixer would change by a factor of $10$ over a dynamic range spanning two orders of magnitude.
Another function of the last stage is to clamp the voltage to be strictly $>0$, in order to prevent an unstable positive feedback behavior, which is done using diode clamp on the output arm~\cite{Horowitz2015}.
Last of all, since the mixer is a current driven device, the last stage of the feedback loop uses TI's OPA627 op-amp which is able to supply tens of mA of current at an output impedance of $50\,\Omega$.

@@

Although it is an essential part of the system, the loop filter is also one of its major limitations. Figure~\ref{fig:noises} shows the noise performance of the output of the fiber amplifiers in open loop as compared to the noise in closed loop. Although noise near DC is lowered (as expected from a closed loop system), noise outside the loop bandwidth is increased in the range of $10$ to $450$ kHz as seen in subfigures (c) and (d).
This can be explained by the Johnson noise~\cite{Horowitz2015} on the $1$ k$\Omega$ input resistors, amplified by the amplification stages of the loop filter.
The sharp drop off at the $450$ kHz point is due to a low-pass filter placed on the output of the loop filter, far above the maximum bandwidth set by the AOM.
When this was determined to be a problem, we decided that a bandwidth of higher than $10\,$ kHz was not required in the foreseeable future.
As a result, we added a first order low pass with a $3$ dB frequency of $10$ kHz leading to the improvement shown in Fig.~\ref{fig:low_pass_noises} (the fiber amplifier used to take this data was different from the one used for~\ref{fig:noises}, although their performance is comparable.

@@

The absolute positions of lattice sites cannot be allowed to change by more than a small fraction ($\approx 10\%$) from experimental shot to shot.
This is ensured via careful design of the optical assembly, since the phase of a retro-reflected lattice is set by the retro-reflecting mirror, which is placed approximately $3\,$cm from the atoms~\cite{Huber2014}.
We have characterized the phase stability of the optical lattice by looking at a long sequence of densely populated images, and found that the phase varies by much less than one lattice site over the course of hours, as seen in Fig.~\ref{fig:averaged_mott}.

@@

The position of the atomic cloud, however, is not determined by the phase of the lattice, but by the underlying harmonic confinement.
In experiments where the entire cloud is addressed by beams auxiliary to the lattice (such as in~\cite{Mazurenko2016} or~\cite{Choi2016}), drift or fluctuation of the lattice position is important.

@@

Over the course of 250 sequential experimental realizations, we monitored the position of the beam on the camera and the position of the atoms.
First, we characterized the passive stability of the optical axis along the transverse $x_t$ and longitudinal $x_l$ directions.
Figure~\ref{fig:stability} a shows a histogram of displacements of atoms between subsequent experimental realizations, showing that the RMS displacement is $0.54(3)$($0.98(4)$) sites along the $x_t$($x_l$) directions.
The difference between displacements along is a geometric effect consistent with the mechanical assembly of the experiment~\cite{Huber2014}.
The distributions are centered within $1\sigma$ of zero along both directions ($-.03(4)$ and $0.01(5)$ sites for transverse and longitudinal, respectively).
If the trap followed a random walk with these step sizes, the experiment would be in trouble, since after only $100$ experimental shots, the trap would move by more than $10$ lattice sites.
The situation is more fortunate, however, since subsequent steps are correlated, as shown in Fig.~\ref{fig:stability} (b).
Once this correlation is taken into account, we estimate a drift of $8(1)\times 10^{-3}\mathrm{sites}/\mathrm{realization}$ in the transverse direction and $12(2)\times 10^{-3}\mathrm{sites}/\mathrm{realization}$ in the axial direction, compared to a beam waist of approximately $80$ $\mu$m ($\approx 140$ sites).

@@

In a real experiment, we cannot directly measure the position of the atoms and must rely on the measurement performed by the camera, related by an affine transformation.
To determine the parameters of the affine transform, we use the remote actuators to purposefully misalign the lattice beams for a small number of shots, and simultaneously measure the atomic positions and the camera reading, then perform a least-squares fit, shown in Fig.~\ref{fig:stability} (c).

Unfortunately, while this fit is acceptable for a short while ($\approx30\mathrm{ shots}$), the quality of the fit degrades over time, correspondingly degrading the predictive power.
The most likely culprit is thermal drift of the camera, since a displacement of only a pixel caused the predicted position to change by approximately $1\,$site, and each camera pixel is about $3\,\mu$m accross.
The camera housing is not engineered for low drift mounting, and the camera heats up by a few degrees Celcius, making a few $\mu$m drift very possible.
To deal with this problem, we point out that the affine transform is composed of three distinct parts - the rotation, scaling and offset.
Of these, only the offset is likely to drift with time.
As a result, we developed a way of updating the calibration over time. To test this procedure, we analyzed the data set of camera measurements and atomic positions in the following way: we split the data set into sequential blocks of $20$ shots, each followed by a block of $8$ shots.
The large blocks were treated as ``masked'' data, which we tried to predict, while the small blocks were treated as ``updating'' events used to update the calibration offsets.
The results are shown in Fig.~\ref{fig:stability} (d-f), showing a large degree of tracking between the predicted positions and measured position - the average displacement between the predicted position and measured position was $0.7\,$lattice sites.
